---
title: "Proyecto Aprendizaje Automatico en la Practica"
author: "Pachamaltese"
date: "May 31, 2015"
output: html_document
---

##Reading and Cleaning Data

Read both training and testing instances. I have used a function named LOAD to load the packages that I will use later.
```{r}
setwd("/Users/pacha/Documents/Coursera/tareas y controles/Practical Machine Learning/proyecto")

load <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
} 

packages <- c("data.table", "Hmisc", "caret", "randomForest", "foreach", "doParallel", "rattle")
load(packages)

training_data <- read.csv("pml-training.csv", na.strings=c("#DIV/0!"," ", "", "NA", "NAs", "NULL"))
evaluation_data <- read.csv("pml-testing.csv", na.strings=c("#DIV/0!"," ", "", "NA", "NAs", "NULL"))
```

I need to drop columns with NAs, drop highly correlated variables and drop variables with 0 (or approx to 0) variance. 

The next chunk will be included to do the calculations but the results are not displayed, or the report would be too long:
```{r, results = 'hide'}
str(training_data)
cleantraining <- training_data[, -which(names(training_data) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window"))]
cleantraining = cleantraining[, colSums(is.na(cleantraining)) == 0] #this drops columns with NAs
zerovariance =nearZeroVar(cleantraining[sapply(cleantraining, is.numeric)], saveMetrics=TRUE)
cleantraining = cleantraining[, zerovariance[, 'nzv'] == 0] #to remove 0 or near to 0 variance variables
correlationmatrix <- cor(na.omit(cleantraining[sapply(cleantraining, is.numeric)]))
dim(correlationmatrix)
correlationmatrixdegreesoffreedom <- expand.grid(row = 1:52, col = 1:52)
correlationmatrixdegreesoffreedom$correlation <- as.vector(correlationmatrix) #this returns the correlation matrix in matrix format
removehighcorrelation <- findCorrelation(correlationmatrix, cutoff = .7, verbose = TRUE)
cleantraining = cleantraining[, -removehighcorrelation] #this removes highly correlated variables (in psychometric theory .7+ correlation is a high correlation)

training_data <- cleantraining

for(i in c(8:ncol(training_data)-1)) {training_data[,i] = as.numeric(as.character(training_data[,i]))}

for(i in c(8:ncol(evaluation_data)-1)) {evaluation_data[,i] = as.numeric(as.character(evaluation_data[,i]))} #Some columns were blank, hence are dropped. I will use a set that only includes complete columns. I also remove user name, timestamps and windows to have a light data set.

feature_set <- colnames(training_data[colSums(is.na(training_data)) == 0])[-(1:7)]
model_data <- training_data[feature_set]
feature_set #now we have the model data built from our feature set.
```
##Cross-Validation

I need to split the sample in two samples. This is to divide training and testing for cross-validation.
```{r}
idx <- createDataPartition(y=model_data$classe, p=0.6, list=FALSE )
training <- model_data[idx,]
testing <- model_data[-idx,]

registerDoParallel()
x <- training[-ncol(training)]
y <- training$classe #After cleaning I use the 70-30 usual proportion of training-testing.
```

##Analysis

Here I define the random forest and the matrix to obtain the answers.
```{r}
set.seed(33) #(can be any number, this is to guarantee reproducibility)
rf <- foreach(ntree=rep(10, 5), .combine=randomForest::combine, .packages='randomForest') %dopar% {
  randomForest(x, y, ntree=ntree) 
}

predictions1 <- predict(rf, newdata=training)
confusionMatrix(predictions1,training$classe)

predictions2 <- predict(rf, newdata=testing)
confusionMatrix(predictions2,testing$classe)
```

##Answers

This was not created by me. This is the function that appeared to submit the answers.
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

x <- evaluation_data
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)

answers

pml_write_files(answers)
```


